Network Configuration Synthesis with Abstract Topologies

### 2016-11-12 paper/abstraction/paper.tex
pass over intro

\newtheorem{thm}{Theorem}[section]
\newtheorem{defn}{Definition}[section]
\newtheorem{lem}[thm]{Lemma}

@@

Computer networks run many critical services, and every second of downtime is
costly at best and dangerous at worst.  Keeping these complex systems up and
running 24/7 is an enormous
challenge~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish,dc-failure-study}.
%
While hardware faults, backhoes, power failures and natural disasters all pose
problems, studies have consistently shown that human error in network
configuration is one of the leading cause of network
outages~\cite{juniper-study,xx}.

To manage and configure large networks today, instead of considering individual
devices, operators classify devices devices into {\em roles}. A role refers to
specific functionality and is played by one or more devices. For instance, in a
data center, roles may correspond to "top-of-rack," "aggregation," and "spine"
routers; and in a backbone network, they may correspond to "core" and "border"
routers. While the network may have hundreds or thousands of devices---a scale
that is impossible for humans to handle---there tend to be a handful of roles.
%
Operators author a configuration {\em template} for each role. Templates are
macros that can be instantiated with different concrete values, based on the
network topology, to generate a configuration for each device.

Unfortunately, template construction is a low-level, manual process. Operators
manually decompose network-wide routing and fault-tolerance policies (e.g.,
prefer one end-to-end path over another or no loss in connectivity after any
one-link failure) across roles and encode them using assembly-language-like
configuration primitives (e.g., role A should not transmit routing information
from B to C). Such decomposition tasks are highly complex~\cite{propane}.

Worse, operators can never be sure that the generated templates correctly
implement the policy,  beyond testing some concrete instantiations using
simulation or verification tools~\cite{x,y}. Testing based on concrete
instantiations is insufficient because large networks evolve continuously, as
devices and links are added or taken offline for maintenance. Templates that
work for the current network topology, may or may not work for future
topologies. Updating the templates, if found to not work for a future topology,
may cause trickle-down, network-wide configuration changes, which is highly
disruptive. Building sophisticated tools for validating templates is difficult
because the templates today have no particular semantics---which policies they
implement and for which topologies---other than the semantics of the
configurations that one obtains from instantiating them.

@@

Given the challenges of generating and validating network configurations, one
might think that operators will be receptive to systems that can generate
provably-correct configurations from high-level policy
specifications~\cite{x,y,z,configassure,propane}. However, our conversations
with two major cloud providers reveal that operators of large networks are
reluctant to use such systems. While they think of their network abstractly, in
terms of roles, these systems require concrete topologies, with no understanding
of roles as a first-order abstraction.  Even if two devices play the same role,
there is no guarantee that the systems will generate (syntactically) similar
configurations; if the operators want to debug or analyze system output, they
will have to consider hundreds of device configurations instead of a handful of
role configurations. These systems become even more brittle when the network
evolves. Any change in network topology requires re-execution of the engine,
from scratch, on the changed topology.  The result may be a completely different
set of configurations.

We present \sysname, a system to generate provably-correct role templates. It
allows operators to input \emph{abstract network topologies} in terms of roles
and their connectivity. For instance, they may specify roles for ``top-of-rack''
and ``aggregation" routers and specify that every top-of-rack router connects to
at least two aggregation routers (to tolerate the failure of a link to an
aggregation router). Other inputs to \sysname include the routing policy of the
network and its fault-tolerance requirements (e.g., the number of link failures
to tolerate). For routing policy, we borrow notation from
Propane~\cite{propane},  which specifies traffic paths at a high-level of
abstraction. But while Propane policies refer to concrete devices, \sysname
policies refer to abstract roles.

Based on these inputs, the \sysname compiler generates one template per role. It
guarantees that, when the templates are instantiated to configurations for {\em
any} concrete topology that complies with the abstract topology, the result
network will correctly implement the routing policy and meet the fault-tolerance
requirements. We show how fault-tolerance analysis can be done over the abstract
topology itself by inferring the worst-case number of edge-disjoint paths in some concrete topology using an optimizing SMT solver.

\sysname templates are also evolution-friendly. When the network migrates from
one concrete topology to another, both of which comply with the abstract
topology, it is guaranteed that only the configurations of devices that acquire
or lose a neighboring device needs to change (e.g., to activate a port on which
the new device is connected).  No other device in the network is impacted.

In addition to creating reusable templates, 
%synthesizing configurations by working directly with an abstract topology means that 
\sysname can scale substantially better than its nearest competitor \propane does: as the number of individual devices in a
network grows, the number of distinct roles will often stay constant.
Consequently, the \sysname compiler can be two orders of magnitude
faster than \propane, taking less than 10 seconds to synthesize templates for large structured networks.

@@

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, the contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item Design of new topology abstractions for network programming
based on graph homomorphisms and connectivity invariants.
\item Design of two new algorithms for analyzing routing policy over
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance
properties of network policy.  Both algorithms provide guarantees for any
concrete network satisfying the abstract connectivity invariants.
\item Design of a template generation strategy that supports network
evolution.
\item Implementation and evaluation of a \sysname compiler that generates
configuration templates and is up to two orders of magnitude faster
than its closest competitor on large networks.
\end{itemize}

@@

\subsection{Template Generation}
\label{subsec:compilation}

@@

%Although \sysname can ensure correct forwarding behavior independent of the exact concrete topology or failure scenario, it is often useful to be able to ask "what if" questions about other network properties. For example, an operator may require a certain level fault tolerance between between different collections of routers. Such analysis depends, not only on the network topology, but also on the underlying routing policy, both of which \sysname has access to.
The \sysname compiler generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} (PG) that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on the \propane compiler, which was designed for concrete networks; we show that it can be extended to abstract inputs and that the extensions are correct.

@@

The PG compactly encodes all paths that protocol messages should be passed along through the network in order to guarantee that the devices find policy-compliant paths dynamically.
%
%Although \sysname policies describe the routes that traffic should take through the network, network protocols (including BGP) typically disseminate router information in the opposite direction. For example, a router Y will tell its neighbor X that it has a route to the destination meaning that X can send traffic through Y. For this reason, we start by compiling
%
Because \sysname policies can have a separate routing policy for each predicate matching traffic, we construct one such PG for each predicate.

@@

The PG is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$.
%
We add edges from the $\mathit{start}$ node to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $P(m)$ denotes the rank of paths through the PG ending at node $m$ and is defined as $P(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\tilde{m} = l$ to extract the topology location from a PG node, when $m = (l, q_{m_1}, \dots, q_{m_k}) \in V'$.

Figure~\ref{fig:example-compilation} shows an example of a PG for the policy that applies to all external traffic from the datacenter example in \S\ref{sec:overview}.
%
%\begin{code}
%\True \Path \Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})
%\end{code}
%\noindent%
%
The first automaton represents the more preferred constraint $\CD{\Exit(N1)}$ and the second automaton represents the less preferred constraint $\CD{\Exit(N2)}$. The PG is shown for both an instance of a simple concrete network matching the abstraction from \S\ref{sec:overview} as well as for the abstract topology.

@@

\begin{defn}
\label{def:homomorphism}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete $PG = (G',\mathit{start},P)$ and abstract $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$ as follows:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}

As we will see in \S\ref{sec:generation}, we have proven that our generation strategy commutes with concrete expansion, meaning that we obtain the same results if we expand the abstraction early, or if we defer the expansion until after template generation.
%
Definition~\ref{def:homomorphism} provides some intuition concerning why our strategy for generating templates is correct.
%Abstract product graphs are related to concrete product graphs via a graph homomorphism, just as abstract topologies are related to concrete topologies via a graph homomorphism.

@@

\section{Template Generation}
\label{sec:generation}

@@

Lemma~\ref{lem:preference} tells us that inferring preferences over the abstract PG before concrete expansion is equivalent to inferring preferences over an already-expanded concrete PG.

@@

\begin{thm}
  For any context $\Gamma$, topologies $G$ and $G^A$, homomorphism $f : G \rightarrow G^A$, and policy $pol$,
  $$\Con(\text{compile}(pol,G^A), \Gamma,f,G) = \text{compile}(\Con(pol,\Gamma,f),G)$$
  \label{thm:concretization}
\end{thm}
%
\vspace{-1em}
Full definitions of compilation and concretization, as well as the proof of Theorem~\ref{thm:concretization} are included in the appendix.

@@

To evaluate \sysname, we apply it on policies for backbone and datacenter networks. We aim to evaluate the expressiveness of our abstracttions, the precision of the abstract disjoint-path analysis, and the generation time of \sysname both with abstraction and without.

@@

\subsection{Generation time}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-time.png}} \\
  \caption{Concrete vs. Abstract Generation Time. \label{fig:compilation-times}}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-analysis-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-analysis-time.png}} \\
  \caption{Abstract Generation Time by Phase. \label{fig:abstract-breakdown}}
  \vspace{-1em}
\end{figure}

Using abstraction allows us to naturally scale network synthesis in \sysname. We evaluate generation time in \sysname both with and without abstraction using routing policy for backbone and datacenter networks inspired by configurations obtained from a large cloud provider. For both types networks, we fix routing policy but scale the size of the underlying topology.

@@

Figure~\ref{fig:compilation-times} shows generation time for the same policies for both the abstract and concrete versions of the policies described above. All experiments are run on an 8 core, 2.4 GHz Intel i7 processor Mac with 8GB of Ram, and the time taken is the average of 3 runs.

For both the networks, the abstract generation starts off slightly slower than concrete generation for small topologies due to the overhead of the fault-tolerance analysis. However, as the size of the topology increases, abstract generation becomes orders of magnitude faster than concrete generation. In all cases for both networks, abstract generation takes less than 10 seconds to complete.

Figure~\ref{fig:abstract-breakdown} shows the relative time taken by each phase of abstract generation. Notably, the fault-tolerance analysis takes the majority of the time, however the analysis does not depend on the number of concrete nodes in the network, and thus is largely a fixed cost. In particular, the number of calls to $\nu$Z remains constant the across topology size. The seesaw behavior for the data center networks results from differences in time taken by $\nu$Z to minimize similar constraints with different values.

>>>>

\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{lem}[thm]{Lemma}

@@

Computer networks run many critical services, and every second of downtime is
costly at best and dangerous at worst.  Yet, keeping these systems
running 24/7 is an enormous
challenge~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish,dc-failure-study}.
%
While hardware faults, backhoes, power failures and natural disasters all pose
problems, studies have consistently shown that human errors in network
configuration is one of the leading cause of 
outages~\cite{juniper-study,xx}.

To configure large networks, instead of considering individual
devices, operators classify devices into {\em roles}. A role refers to
specific functionality and is served by one or more devices. For instance, in a
data center, roles may be "top-of-rack," "aggregation," and "spine"
routers; and in a backbone network, they may be "core" and "border"
routers. While the network may have hundreds or thousands of devices---a scale
that is impossible for humans to handle---there tend to be a handful of roles.
%
Operators author a configuration {\em template} for each role. Templates are
macros that, given a network topology, can be instantiated with different concrete values to generate device configurations.

Unfortunately, template construction is a low-level, manual process. Operators
manually decompose network-wide routing and fault-tolerance policies (e.g.,
prefer one path over another, or a link failure should not cause connectivity loss) across different roles and encode the decomposition using assembly-language-like
configuration primitives (e.g., role A should not transmit routing information
from B to C). Such decomposition and encoding is highly complex~\cite{propane}.

Worse, operators can never be sure that their templates correctly implement the policy. %
Formally verifying them is difficult; today's hand-generated templates have no particular semantics---which policies they implement and for which topologies---other than those of the configurations obtained from instantiating them. 
%
Thus, operators validate templates by testing some concrete instantiations using
simulation or verification tools~\cite{x,y}. This testing is insufficient because large networks evolve continuously, as
devices and links are added or taken offline for maintenance. Templates that
work for the current network topology, may or may not work for future
topologies, and it is intractable to test all possible future topologies. Updating the templates, if found to not work for a future topology,
may cause trickle-down, network-wide configuration changes, which is highly
disruptive. 
%Building sophisticated tools for validating templates is difficult

@@

Given the challenges of generating and validating network configurations, one
might think that operators will be receptive to systems that can generate
provably-correct configurations from high-level policy
specifications~\cite{x,y,z,configassure,propane}. However, our conversations
with two major cloud providers reveal that operators of large networks are
reluctant to use such systems. While they think of their network abstractly, in
terms of roles, these systems only understand concrete topologies, with no understanding of roles.  Even if two devices play the same role,
there is no guarantee that the systems will generate (syntactically) similar
configurations; if the operators want to debug or analyze system output, they
will have to consider hundreds of device configurations instead of a handful of
role configurations. These systems become even more brittle when the network
evolves. Any change in network topology requires re-execution of the engine,
from scratch, on the new topology.  The result may be a completely different
set of configurations.

We present \sysname, a system to generate provably-correct role templates. It
allows operators to input \emph{abstract network topologies} in terms of roles
and their connectivity. For instance, they may specify roles for ``top-of-rack''
and ``aggregation" routers and specify that every top-of-rack router connects to
at least two aggregation routers (to tolerate the failure of a link to an
aggregation router). Other inputs to \sysname include the routing policy of the
network and its fault-tolerance requirements (e.g., the number of link failures
to tolerate). For high-level specification of routing policy, we borrow notation from
Propane~\cite{propane}. But while Propane policies refer to concrete devices, \sysname
policies refer to abstract roles.

Based on these inputs, the \sysname compiler generates one template per role. It
guarantees that, when the templates are instantiated to configurations for {\em
any} concrete topology that complies with the abstract topology, the result
network will correctly implement the routing policy and meet the fault-tolerance
requirements. We show how precise fault-tolerance analysis can be done in the abstract domain by using a set of sound inference rules to infer the lower bound on the number of edge-disjoint paths in any concrete topology. 
%Our analysis uses a set of sound inference rules and  using an optimizing SMT solver.

\sysname's abstract reasoning provides two other benefits. First, its templates are evolution friendly. When the network migrates from
one concrete topology to another, both of which comply with the abstract
topology, only the configurations of devices that acquire or lose a neighbor need to change. This guarantee is the best that any system can give as neighbor relationships are explicitly configured in devices. 
%No other device in the network is impacted.

Second, \sysname can scale substantially better than its nearest competitor \propane that operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgeries-popl16}.
Consequently, the \sysname compiler can be two orders of magnitude
faster than \propane, taking less than 10 seconds to synthesize templates for large structured networks.

@@

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, our contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item Design of new topology abstractions for network programming
based on graph homomorphisms and connectivity invariants.
\item Design of two new algorithms for analyzing routing policy over
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance
properties of network policy.  Both algorithms provide guarantees for any
concrete network satisfying the abstract connectivity invariants.
%\item Design of a template generation strategy that supports network
%evolution.
\item Implementation and evaluation of the \sysname compiler that generates
provably-correct templates and is much faster
than a compiler based on concrete topologies.
\end{itemize}

@@

\subsection{Compilation}
\label{subsec:compilation}

@@

%Although \sysname can ensure correct forwarding behavior independent of the exact concrete topology or failure scenario, it is often useful to be able to ask "what if" questions about other network properties. For example, an operator may require a certain level fault tolerance between between different collections of routers. Such analysis depends, not only on the network topology, but also on the underlying routing policy, both of which \sysname has access to.
The \sysname compiler generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} (PG) that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on the Propane compiler, which was designed for concrete topologies and policy; we show that it can be extended to abstract inputs and that the extensions are correct.

@@

The PG compactly encodes all paths protocol messages should be passed along through the network in order to guarantee the devices find policy-compliant paths dynamically.
%
%Although \sysname policies describe the routes that traffic should take through the network, network protocols (including BGP) typically disseminate router information in the opposite direction. For example, a router Y will tell its neighbor X that it has a route to the destination meaning that X can send traffic through Y. For this reason, we start by compiling
%
Because \sysname policies can have a separate routing policy for each predicate matching traffic, we construct one such PG for each predicate.

@@

The PG is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$ is a valid topology link.
%
We add edges from the start node to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $P(m)$ denotes the rank of paths through the PG ending at node $m$ and is defined as $P(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\tilde{m} = l$ to extract the topology location from a PG node, when $m = (l, q_{m_1}, \dots, q_{m_k}) \in V'$.

Figure~\ref{fig:example-compilation} shows an example of a PG for the policy that applies to all external traffic from the datacenter example in \S\ref{sec:overview} ($\True ~\Path~ \CD{\Exit(N1 \Prefer~N2)}$).
%
%\begin{code}
%\True \Path \Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})
%\end{code}
%\noindent%
%
The first automaton represents the more preferred constraint $\CD{\Exit(N1)}$ and the second automaton represents the less preferred constraint $\CD{\Exit(N2)}$. The PG is shown for both an instance of a simple concrete network matching the abstraction from \S\ref{sec:overview} as well as for the abstract topology.

@@

\begin{defn}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete $PG = (G',\mathit{start},P)$ and abstract $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$ as follows:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}

@@

\section{Template Generation}

@@

Lemma~\ref{lem:preference} gives us reason to believe that we can perform compilation over the abstract PG directly and ensure that the inferred device-level preferences will be correct when instantiating the template with a concrete network.

@@

\begin{thm}
  For any context $\Gamma$, topology $G$, abstract topology $G^A$, homomorphism $f : G \rightarrow G^A$, and policy $pol$,

  \noindent
  $\Con(\text{compile}(pol,G^A), \Gamma,f,G) = \text{compile}(\Con(pol,\Gamma,f),G)$
  \label{thm:concretization}
\end{thm}
%
\vspace{-1em}
Full definitions of compilation and concretization, as well as the proof of Theorem~\ref{thm:concretization} are included in the appendix.

@@

To evaluate \sysname, we apply it on policies for backbone and datacenter networks. We aim to evaluate the expressiveness of our abstracttions, the precision of the abstract disjoint-path analysis, and the compilation time of \sysname both with abstraction and without.

@@

\subsection{Compilation time}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-time.png}} \\
  \caption{Concrete vs. Abstract Compilation Time. \label{fig:compilation-times}}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-analysis-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-analysis-time.png}} \\
  \caption{Abstract Compilation Time by Phase. \label{fig:abstract-breakdown}}
  \vspace{-1em}
\end{figure}

Using abstraction allows us to naturally scale network synthesis in \sysname. We evaluate compilation time in \sysname both with and without abstraction using routing policy for backbone and datacenter networks inspired by configurations obtained from a large cloud provider. For both types networks, we fix routing policy but scale the size of the underlying topology.

@@

Figure~\ref{fig:compilation-times} shows compilation time for the same policies for both the abstract and concrete versions of the policies described above. All experiments are run on an 8 core, 2.4 GHz Intel i7 processor Mac with 8GB of Ram, and the time taken is the average of 3 runs.

For both the networks, the abstract compilation starts off slightly slower than concrete compilation for small topologies due to the overhead of the fault-tolerance analysis. However, as the size of the topology increases, abstract compilation becomes orders of magnitude faster than concrete compilation. In all cases for both networks, abstract compilation takes less than 10 seconds to complete.

Figure~\ref{fig:abstract-breakdown} shows the relative time taken by each phase of abstract compilation. Notably, the fault-tolerance analysis takes the majority of the time, however the analysis does not depend on the number of concrete nodes in the network, and thus is largely a fixed cost. In particular, the number of calls to $\nu$Z remains constant the across topology size. The seesaw behavior for the data center networks results from differences in time taken by $\nu$Z to minimize similar constraints with different values.
